An equivalence of compositional frameworks

### 2018-10-06 EquivCompFrameworks_20180911kc.tex
minor changes, cleaning up and adding to applications section

Let $\mathrm{A}$ be a category with finite colimits and $F \colon \mathrm{A} \to \bold{Cat}$ a lax monoidal pseudofunctor. Then there exists a double category $\mathbb{F} \textnormal{Cospan}(\mathrm{A}) \coloneqq (\mathbb{G}/ \mathbb{F})$ which has:

@@

The unit structure functor $U \colon (\mathbb{G},\mathbb{F})_0 \to (\mathbb{G},\mathbb{F})_1$ is defined on objects as: 

@@

together with the morphism $\iota_{!_f} = F(f) F(!) \phi \colon 1 \to F(c^\prime)$. We also have source and target structure functors $s,t \colon (\mathbb{G} / \mathbb{F})_1 \to (\mathbb{G} / \mathbb{F})_0$ where the source of a horizontal 1-cell

@@

In this section we prove that two different frameworks which utilize symmetric monoidal double categories, namely decorated cospans and `structured cospans', are equivalent under the conditions required from the previous section that allow for a left adjoint $L \colon \mathrm{A} \to \mathrm{X}$ to be obtained from a symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \bold{Cat}$. The double category version of decorataed cospans is given in Section \ref{DecCospansDoublecat}. The other framework which also utilizes symmetric monoidal double categories, due to Baez and the first author and goes by the name of `structured cospans', is explained below. Once again following the notation of Shulman \cite{Shul2}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:

@@

@@

\textbf{Also talk about electrical circuit, Markov process and Petri net examples.}

>>>>

Let $\mathrm{A}$ be a category with finite colimits and $F \colon \mathrm{A} \to \bold{Cat}$ a lax monoidal pseudofunctor. Then there exists a double category $\mathbb{F} \textnormal{Cospan}(\mathrm{A})$ which has:

@@

The unit structure functor $U \colon \mathbb{F}\textnormal{Cospan}(\mathrm{A})_0 \to \mathbb{F}\textnormal{Cospan}(\mathrm{A})_1$ is defined on objects as: 

@@

together with the morphism $\iota_{!_f} = F(f) F(!) \phi \colon 1 \to F(c^\prime)$. We also have source and target structure functors $S, T \colon \mathbb{F}\textnormal{Cospan}(\mathrm{A})_1 \to \mathbb{F}\textnormal{Cospan}(\mathrm{A})_0$ where the source of a horizontal 1-cell

@@

In this section we prove that two different frameworks which utilize symmetric monoidal double categories, namely decorated cospans and `structured cospans', are equivalent under the conditions required from the previous section that allow for a left adjoint $L \colon \mathrm{A} \to \mathrm{X}$ to be obtained from a symmetric lax monoidal pseudofunctor $F \colon \mathrm{A} \to \bold{Cat}$. The double category version of decorataed cospans is given in Section \ref{DecCospansDoublecat}. The other framework also utilizing symmetric monoidal double categories due to the first two authors and which goes by the name of `structured cospans', is explained below. Once again following the notation of Shulman \cite{Shul2}, given a double category $\mathbb{A}$, we write $_f \mathbb{A}_g(M,N)$ for the set of 2-morphisms in $\mathbb{A}$ of the form:

@@

In this section we present several examples each of which may be realized in the context of decorated cospans or in the context of structured cospans. The first example regarding graphs was mentioned in the introduction and used as a reoccurring theme throughout the paper. The next three examples which take on more of an applied flavor, consists of electrical circuits, Markov processes and Petri nets. Each of these has been studied extensively by the first author and others by way of `black-boxing'. Black-boxing is a way of interpreting the behavior of an open system, that is, a system with prescribed inputs and outputs such as the terminals of an electrical circuit, by observing the activity at the inputs and the outputs. The semantics of the activity at an open system's inputs and outputs is typically described in a category such as $\mathrm{LinRel}$ of finite dimensonal vector spaces and linear relations. Thus, in each case, black-boxing results in functors such as: $$\blacksquare_1 \colon \mathrm{Circ} \to \mathrm{LinRel}$$ $$\blacksquare_2 \colon \mathrm{Mark} \to \mathrm{LinRel}$$ $$\blacksquare_3 \colon \mathrm{Petri} \to \mathrm{LinRel}.$$ Each of these black-boxing functors also possess other convenient properties such as being symmetric monoidal. The first two of these were first done using Fong's theory of decorated cospans and then extended by the first two authors using structured cospans. The last two of these were also extended by being realized as double functors between double categories. \textbf{Make this better...}
\subsection{Graphs}

@@

\subsection{Electrial circuits}

\subsection{Markov processes}

\subsection{Petri nets}
